---
title: "Cruise Ships' Impact on NYC Air Quality"
subtitle: "Final Presentation, DATA 698 CUNY SPS"
author: "Alice Friedman"
institute: "CUNY School for Professional Studies"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(forecast)
library(scales)
library(ggcorrplot)
library(httr)
library(jsonlite)
library(tidyr)
library(forcats)
library(glmnet)
library(gridExtra)
library(stringr)
```

## Introduction

- Air Pollution is responsible for an estimated 6.7 million annual premature deaths globally (WHO, 2022).
  
- Cruise ship emissions are a visible source of urban pollition in coastal cities, with increasing attention to mitigation strategies like the use of on-shore power while in port

- *Research Gap*: Lack of empirical analysis in the context of New York City, emphasizing the need for a data-driven approach.

## Study Area: Brooklyn Cruise Terminal

- Only terminal on the east coast with shore power system designed to reduce emissions

- Installed in 2016 at the cost of $21M

- Red Hook neighborhood has worst air quality in Brooklyn (NYC.gov, 2023)

- Policy Context: New York City considering local law Int. 1050, requiring cruise ships to use on-shore power by 2028 at all ports (including Manattan Cruise Terminal)

## Research Questions

-  What level of expected improvement in air quality can be expected as a result of improvements to New York City cruise ship terminals and proposed new regulations that will require all cruise ships to use on-shore power by 2028?

-  What feature set among available inputs will best predict air quality *in the absence of consideration of cruise ship inputs*? 

- Does the inclusion of cruise ship features measurably improve these results?

## Literature Review

- Selection of PM 2.5 as a critical parameter due to its known health impacts and reliable measurement methods

- Methodological research included use of circular data (wind direction) and PM 2.5 readings from Purple Air

- Studies from Europe (Hrvoje, 2011; Ruiz-Guerra et al, 2019; Fabregat et al, 2021) and Canada (Poplawski et al, 2011) support cruise ship emissions as measurably impacting on-shore air quality, typically using linear regression

- No studies found to document this relationship in New York City, or using non-parametric methods to identify feature set

## Data Sources

- Weather and airquality data at the Port is taken from a Dyson Instruments weather station located adjacent to the Brooklyn Cruise Terminal

- Observations are taken in 15 minute increments, starting July 11, 2023

- Ship docking data is taken from Port of New York, hourly, and the EDC for information on whether ships can access the on-shore power or must use the generator (all ships outside of BCT use generators)

- Additional data taken from Purple Air sensor at Dyberry farm as a control for wildfire smoke

## Data Preparation

- Weather data features evaluated include temperature, dew point, heat bulb index, humidity, wind speed, wind direction, air pressure reported at average, low, and high as well as rain in inches

- Highly correlated weather features pruned to include only average value

- Wind direction converted to radians and then defined by the sine and cosine of theta to avoid correlation

- Cruise ship compatibility is taken from the EDC and extrapolated to ship dockings in July

- PM 2.5 levels from Purple Air corrected based on EPA 

- Weather data aggregated hourly (average value, except for rain which is summed) to join with cruise ship data and control sensor data from Purple Air

- Final data set has 17 features, 5 possible targets, and 3,396 observations across 141 days (some days contain more than 24 observations because there may be more than one ship present)

## Time Series

```{r load_data, message=FALSE, warning=FALSE}
#source("purple_pull.R") pulls from API and writes to csvs
source("cruise_dataprep.R") #loads CSV generated from worlshipny.org and merges with EDC info on shore power
source("PortSide_dataprep.R") #loads CSV aggregates as hourly data ("Dyson_hourly") and daily"
```

```{r join_data_hourly}
Dyberry <- read_csv("Dyberry.csv", show_col_types = FALSE ) #loads PM2.5 from purple air at Dyberry Farm

ship_cols <-c(
  "Ship_Presence","Terminal", "Line", "Generator"
)

drop_list <-c(
  "degrees_mean",
  "Arrival", "ArrivalHour",
  "Departure", "DepartureHour",
  "degrees_mean"
)

#Join PortSide NY data to purple air data to cruise ships data to create an hourly data frame
df_hourly <- left_join(Dyson_hourly, dplyr::select(Dyberry, local_datetime, PM2.5_corrected), by=c("hour" = "local_datetime"))  %>%     left_join(cruise_ships_hourly, by=c("hour"="Hour")) %>%
  rename(
    PM2.5_Dyson = PM2.5_mean
    ) %>%
  relocate(hour, all_of(ship_cols)) %>% 
  dplyr::select(-all_of(drop_list)) %>%
  mutate( # add features related to hour and day of week
    dayofweek = wday(hour),
    hourofday = hour(hour)
  )
write_csv(df_hourly, "Hourly_Joined.csv")
```

```{r join_daily}
df_daily<- Dyson_daily %>% left_join(cruise_ships_daily, by=c("date"="Date")) %>%
  relocate(date, any_of(ship_cols)) %>% 
  dplyr::select(-any_of(drop_list)) %>%
  mutate(
    dayofweek = wday(date)
    ) 

write_csv(df_daily, "Daily_Joined.csv")
```

```{r plot1}
p_time_PM <- ggplot(df_hourly %>% filter(complete.cases(.))) +
  geom_line(aes(x=hour, y=PM2.5_Dyson, color="PM 2.5 Levels in Red Hook")) +
  geom_line(aes(x=hour, y=PM2.5_corrected, color="PM 2.5 Levels Upstate\n(Dyberry Farm Control)")) +
  scale_x_datetime(breaks = scales::date_breaks("1 month"), labels = date_format("%m %d")) +
  ylab("PM 2.5 (Parts Per Million)") + 
  xlab("Date") +
  theme(aspect.ratio = 1/4)+ 
  scale_color_manual(values = c("blue", "lightblue"), name = "Legend Title")

p_time_AQI <- ggplot(Dyson) +
  geom_line(aes(x=DateTime, y=AQI, color = "Actual AQI")) +  
  geom_line(aes(x=DateTime, y=AQI_adj, color = "Adjusted for Woodsmoke")) +  
  scale_x_datetime(breaks = scales::date_breaks("1 month"), labels = date_format("%m %d")) +
  ylab("AQI") + 
  xlab("Date") +
  theme(aspect.ratio = 1/4)+ 
  scale_color_manual(values = c("blue", "lightblue"), name = "Legend Title")

grid.arrange(p_time_AQI, p_time_PM, ncol=1)
```






